Blueprint for efficient nuclear spin characterization in color center

### 2023-08-23 article.tex
a

\author{Majid Zahedian}
 \affiliation{3rd physical institute University of Stuttgart, Germany}
  \affiliation{Max Planck School of Photonics, Germany}%Lines b%Lines break automatically or can be forced with \\

@@

is necessary for enlarging the quantum register and precise description of the electron spin dynamics. Traditionally, nuclear spin characterisation is done with Optically Detected Magnetic Resonance (ODMR) (ref to Phillip Neumann's works), however the spectroscopic resolution is limited to the $T_2^\ast$ of the electron spin. Hence, some sequence is needed to increase the resolution. People have used Hahn Echo sequence to refocus the electron spin and enhance the coherence time of the electron spin \cite{childress2006coherent}. But this method only works for defects with a specific nuclear spins configuration. Depending on the applied protocol for nuclear spins characterization, the number of accessible nuclear spin can increase. Electron Spin Echo Envelop Modulation (ESEEM) types of sequences allow us to access the highest number of nuclear spins since the spectroscopy resolution is limited to the longitudinal relaxation time ($T_1$) of the system (see fig. \ref{fig:1}). 
\begin{figure*}%[htbp]
	\begin{center}
		\includegraphics[width=0.7\textwidth]{pict/drawing0.pdf}
		\caption{a. nuclear spins that can be characterized using each sequence. b. number of nuclear spins as a function of distance for the defect. The dashed line shows having T1 = 1s limitation, how many nuclear spins are accessible in diamond. }
		\label{fig:1}
	\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%================= %THEORY
\section{Methods}
\label{sec:theory}
%%%%%%%%%%%%%%%Í%%%%=================
In this section, we provide the theoretical formulas for. In a central electron spin with the electron spin $\boldmath{S}$, we only apply pulses resonant with two of the electron spin sublevels, denoted as $m_s=s_0$ and $m_s = s_1$. These two spin sublevels are separated, due to Zeeman effect or zero field splitting, and form an artificial two-level system with energy splitting of $\omega_a$. Now, we add spin $\frac{1}{2}$nuclear spins to the system (for considering P1 centers, one has to add nuclear quadrupole to the presented Hamiltonian.). The Hamiltonian of a central electron spins system surrounded by nuclear spins in the secular approximation in the rotating frame of the applied microwave $\omega_{mw}$ (considering rotating wave approximation) can be written as:
\begin{align}
	H = \Delta S_z + \sum_k (\omega_L^{(k)}  + A_{zz}^{(k)} S_z) I_z^{(k)} + A_{zx}^{(k)} S_z I_x^{(k)}
	\label{eq:H}
\end{align}
Where $\Delta = \omega_a - \omega_{mw}$ is the detuning, $\omega_L^{(k)} =\gamma_n^{(k)} B$ is nuclear Rabi frequency with $\gamma_n^{(k)}$ nuclear spin gyromagnetic ratio and $B$ external magnetic field, $A_{zz}, A_{zx}$ parallel and perpendicular secular component of hyperfine tensor. This Hamiltonian results that depending on the electron spin sublevel $s_i$ ($i=0,1$), the nuclear spin rotate with the resonance frequency of $\omega_i = \sqrt{(\omega_L + s_i A_{zz})^2+ (s_i A_{zx})^2}$, $i=0,1$ along the axes $\vec{n}_i =(\frac{s_i A_{zx}}{\omega_i}, 0, \frac{\omega_L + s_i A_{zz}} {\omega_i})$. It is assumed that the pulse duration $t_p$ are short enough ($\omega_L t_p \ll 1$) that the nuclear spin dynamics during this time is negligible. The most simplest characterization sequence is Ramsey. \\ The Ramsey signal can be obtained as follows:
	\expval{\sigma_z}_{Ram} = \cos(\Delta \tau) \prod_{k=1}^{n} \Big[ \cos(\frac{\omega_0 \tau}{2}) \cos(\frac{\omega_1 \tau}{2})+ \notag \\
	(\vec{n_0}\cdot \vec{n_1})\sin(\frac{\omega_0 \tau}{2}) \sin(\frac{\omega_1 \tau}{2}) \Big]^{(k)}
	\textcolor{gray}{e^{-\frac{\tau}{T_2^\ast}}}
We do not take any relaxation process into account and the gray exponential decay is put by hand to remember the relaxation time scale. Ramsey signal decay with electron spin $T_2^\ast$, so it can be used to observe the core of the defect. Ramsey sequence is sensitive to detuning since the electron spin is not refocused, however it is possible to sense nuclear spins with vanishing perpendicular hyperfine coupling.\\
\begin{figure}%[htbp]
		\includegraphics[width=0.9\columnwidth]{pict/seqs.png}
		\caption{Pulse sequences a. Ramsey b. Hahn Echo c. Dynamical Decoupling d. DDESEEM}
		\label{fig:seqs}
\end{figure} 

@@

\begin{figure*}[htbp]

@@

		\includegraphics[width=0.8\textwidth]{pict/fig3.png}
		\caption{a. DD for NV in diamond b. DD for group IV defect in diamond }

@@

\begin{figure*}[htbp]
	\begin{center}
		\includegraphics[width=0.8\textwidth]{pict/Fisher.pdf}
		\caption{a. DD spin 1 b. DD spin 1/2 c. 5pESEEM spin 1 d. 5pESEEM spin 1/2 }
		\label{fig:Fisher}
	\end{center}
\end{figure*}

We want to make a sequence that is limited by T1 instead of T2. The idea of 5-pulse ESEEM \cite{kasumaj20085} is that to create an entanglement (Hahn-echo) before the free evolution (Ramsey). In other words, change the initial density matrix for the Ramsey sequence from such that entanglement already exist in the electron polarization terms of the density matrix. Fig \ref{fig:seqs} shows the sequence, which can be interpreted as two Hahn Echos that are separated by a long free evolution $T\gg T_2^\ast$ such that the coherence of the electron spin vanishes. The analytical formula to this sequence is provided in the appendix \ref{app:5pESEEM}. During the free evolution time, each nuclear spins oscillates with either of two resonance frequencies obtained from eq. \ref{eq:H}.
\begin{figure*}[htbp]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{pict/fig4.png}
		\caption{a. FFT of ESEEM signal b. a zoom into weakly coupled nuclear spins c. Enhancing weakly coupled nuclear spins by having 72 pulses in each entangling period. $\tau$ is set to the Larmor bright spot.}
		\label{fig:4}
	\end{center}
\end{figure*}
Fig \ref{fig:4} shows the furrier transform of the signal. Zooming into area close to the Larmor frequency reveals the weakly coupled nuclei. Each nuclear spins have two peaks in the spectrum. Expanding the resonance frequencies up to first order with respect to hyperfine coupling results that the weakly coupled nuclei keep their order as they go further away from the nuclei. This fact makes their characterization rather straight forward. To observe the weakly coupled nuclei more pronouncedly, one can use DDESEEM sequence by adding more $\pi$ pulses in the entangling periods. To distinguish other nuclei, extra measurements is required. One method is to use two dimensional hyperfine correlation spectroscopy (2D Hyscore)  \cite{vorobyov2022addressing}. However, this method is rather time-consuming as two time variables has to be swept. We suggest sweeping $\tau$ parameter and keep the track of frequency amplitude as a function of $\tau$. Two frequencies that belong to the same nuclear are correlated because the modulation depth oscillates with blind spot terms of both frequencies $\sin^2(\frac{\omega_0 \tau}{2}) \sin^2(\frac{\omega_1 \tau}{2})$. Fig. \ref{fig:5} shows two frequencies goes to bright and blind spots simultaneously. Hence, taking two dimensional correlation of the spectrum, one can deduce which two frequencies are correlated and belong to the same nuclear. \\

>>>>

\author{Majid Zahedian}
 \affiliation{3rd physical institute University of Stuttgart, Germany}
  %\affiliation{Max Planck School of Photonics, Germany}%Lines b%Lines break automatically or can be forced with \\

@@

is necessary for enlarging the quantum register and precise description of the electron spin dynamics. Traditionally, nuclear spin characterisation is done with Optically Detected Magnetic Resonance (ODMR) (ref to Phillip Neumann's works), however the spectroscopic resolution is limited to the $T_2^\ast$ of the electron spin. Hence, some sequence is needed to increase the resolution. People have used Hahn Echo sequence to refocus the electron spin and enhance the coherence time of the electron spin \cite{childress2006coherent}. But this method only works for defects with a specific nuclear spins configuration. Depending on the applied protocol for nuclear spins characterization, the number of accessible nuclear spin can increase. Electron Spin Echo Envelop Modulation (ESEEM) types of sequences allow us to access the highest number of nuclear spins since the spectroscopy resolution is limited to the longitudinal relaxation time ($T_1$) of the system (see fig. \ref{fig:1}). 
\begin{figure*}%[htbp]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{pict/drawing0.pdf}
		\caption{Comparing DD and 5pESEEM as two nuclear spin characterization method in different systems. a. An electron spin one system: in DD signal, each nuclear spin has a distinct resonance time up to first order of $A_{zz}$; the accuracy of the obtained hyperfine interaction is limited to $T_2$ of the electron spin. In the Fourier transform of the ESEEM signal, the two resonance frequency of each spin is present; the accuracy of the obtained hyperfine interaction is limited to $T_1$ of the electron spin b. An electron spin one system with two strongly coupled nuclear spins and many weakly coupled spins. No resonance time is detectable since the required condition for DD ($\omega_L \gg A_{zz}, A{zx}$) does not hold. However, resonant frequencies is still obtainable in the ESEEM signal c. An electron spin one half system: nuclear spins have more or less the same resonance time as it depends on the second order of hyperfine coupling. However, resonant frequency of each spins is distinguishable in the ESEEM signal. d. A system containing two different nuclear spin species: two nuclear spin bath can interfere in the DD signal, which makes the signal analysis more challenging. However, the species can be inferred from ESEEM signal since spe cies have different Larmor frequencies.}
		\label{fig:1}
	\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%================= %THEORY
\section{Methods}
\label{sec:theory}
%%%%%%%%%%%%%%%Í%%%%=================
In this section, we provide the theoretical formulas for. In a central electron spin with the electron spin $\boldmath{S}$, we only apply pulses resonant with two of the electron spin sublevels, denoted as $m_s=s_0$ and $m_s = s_1$. These two spin sublevels are separated, due to Zeeman effect or zero field splitting, and form an artificial two-level system with energy splitting of $\omega_a$. Now, we add spin $\frac{1}{2}$nuclear spins to the system (for considering P1 centers, one has to add nuclear quadrupole to the presented Hamiltonian.). The Hamiltonian of a central electron spins system surrounded by nuclear spins in the secular approximation in the rotating frame of the applied microwave $\omega_{mw}$ (considering rotating wave approximation) can be written as:
\begin{align}
	H = \Delta S_z + \sum_k (\omega_L^{(k)}  + A_{zz}^{(k)} S_z) I_z^{(k)} + A_{zx}^{(k)} S_z I_x^{(k)}
	\label{eq:H}
\end{align}
Where $\Delta = \omega_a - \omega_{mw}$ is the detuning, $\omega_L^{(k)} =\gamma_n^{(k)} B$ is nuclear Rabi frequency with $\gamma_n^{(k)}$ nuclear spin gyromagnetic ratio and $B$ external magnetic field, $A_{zz}, A_{zx}$ parallel and perpendicular secular component of hyperfine tensor. This Hamiltonian results that depending on the electron spin sublevel $s_i$ ($i=0,1$), the nuclear spin rotate with the resonance frequency of $\omega_i = \sqrt{(\omega_L + s_i A_{zz})^2+ (s_i A_{zx})^2}$, $i=0,1$ along the axes $\vec{n}_i =(\frac{s_i A_{zx}}{\omega_i}, 0, \frac{\omega_L + s_i A_{zz}} {\omega_i})$. It is assumed that the pulse duration $t_p$ are short enough ($\omega_L t_p \ll 1$) that the nuclear spin dynamics during this time is negligible. The most simplest characterization sequence is Ramsey. \\ The Ramsey signal can be obtained as follows:
\begin{gather}
	\expval{\sigma_z}_{Ram} = \cos(\Delta \tau) \prod_{k=1}^{n} \Big[ \cos(\frac{\omega_0 \tau}{2}) \cos(\frac{\omega_1 \tau}{2})+ \notag \\
	(\vec{n_0}\cdot \vec{n_1})\sin(\frac{\omega_0 \tau}{2}) \sin(\frac{\omega_1 \tau}{2}) \Big]^{(k)}
	\textcolor{gray}{e^{-\frac{\tau}{T_2^\ast}}}
\end{gather} 
We do not take any relaxation process into account and the gray exponential decay is put by hand to remember the relaxation time scale. Ramsey signal decay with electron spin $T_2^\ast$, so it can be used to observe the core of the defect. Ramsey sequence is sensitive to detuning since the electron spin is not refocused, however it is possible to sense nuclear spins with vanishing perpendicular hyperfine coupling.\\

@@

\begin{figure*}%[htbp]

@@

		\includegraphics[width=0.9\textwidth]{pict/dd_compare2.pdf}
		\caption{Simulated DD signal with 64 $\pi$-pulses for the register reported \cite{abobeih2019atomic} assuming the electron spin is a. spin one b. spin one-half system. Simulated Fisher information for the case of the electron spin c. one d. one-half system.}

@@

We want to make a sequence that is limited by T1 instead of T2. The idea of 5-pulse ESEEM \cite{kasumaj20085} is that to create an entanglement (Hahn-echo) before the free evolution (Ramsey). In other words, change the initial density matrix for the Ramsey sequence from such that entanglement already exist in the electron polarization terms of the density matrix. Fig \ref{fig:seqs} shows the sequence, which can be interpreted as two Hahn Echos that are separated by a long free evolution $T\gg T_2^\ast$ such that the coherence of the electron spin vanishes. The analytical formula to this sequence is provided in the appendix \ref{app:5pESEEM}. During the free evolution time, each nuclear spins oscillates with either of two resonance frequencies obtained from eq. \ref{eq:H}.
\begin{figure*}[htbp]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{pict/eseem_compare2.pdf}
		\caption{Simulated FFT of ESEEM signal with a zoom into weakly coupled nuclear spins for the register reported \cite{abobeih2019atomic} assuming the electron spin is one-half. b. Enhancing the sensitivity to weakly coupled spins by increasing number of $\pi$-pulses to $N=72$ pulses in each entangling period. $\tau$ is set to the Larmor bright spot. Simulated Fisher information for the case of the electron spin c. one d. one-half system.}
		\label{fig:4}
	\end{center}
\end{figure*}
Fig \ref{fig:4} shows the furrier transform of the signal. Zooming into area close to the Larmor frequency reveals the weakly coupled nuclei. Each nuclear spins have two peaks in the spectrum. Expanding the resonance frequencies up to first order with respect to hyperfine coupling results that the weakly coupled nuclei keep their order as they go further away from the nuclei. This fact makes their characterization rather straight forward. To observe the weakly coupled nuclei more pronouncedly, one can use DDESEEM sequence by adding more $\pi$ pulses in the entangling periods. To distinguish other nuclei, extra measurements is required. One method is to use two dimensional hyperfine correlation spectroscopy (2D Hyscore)  \cite{vorobyov2022addressing}. However, this method is rather time-consuming as two time variables has to be swept. We suggest sweeping $\tau$ parameter and keep the track of frequency amplitude as a function of $\tau$. Two frequencies that belong to the same nuclear are correlated because the modulation depth oscillates with blind spot terms of both frequencies $\sin^2(\frac{\omega_0 \tau}{2}) \sin^2(\frac{\omega_1 \tau}{2})$. Fig. \ref{fig:5} shows two frequencies goes to bright and blind spots simultaneously. Hence, taking two dimensional correlation of the spectrum, one can deduce which two frequencies are correlated and belong to the same nuclear. \\
